Sequential Composition for Relaxed Memory: Pomsets with Predicate Transformers

### 2021-07-07 paper/intro.tex
Adding text to introduction to plumb in tool

In certain cases it is possible to ban concurrent access using separation
\cite{OHearn:2007:RCL:1235896.1236121}, or to accept inefficient
implementation in order to obtain sequential consistency
\cite{DBLP:conf/snapl/MarinoMMNS15}.  When these approaches are not
available, however, we are left with an enormous gap in our understanding of
one of the most basic elements of computing: the humble semicolon.  Until
recently, existing approaches either
\begin{itemize}
\item did not bother tracking dependencies, allowing ``thin air'' executions
  \cite{Batty:2011:MCC:1926385.1926394},
\item tracked dependencies conservatively, using syntax, requiring inefficient
  implementation of relaxed access
  \cite{DBLP:conf/pldi/LahavVKHD17,Boehm:2014:OGA:2618128.2618134,DBLP:conf/oopsla/VafeiadisN13,DBLP:journals/corr/abs-1804-04214}---while
  this may be fine for C, it's a non-started for safe languages like Java,
  which lack C's ``plain'' access,
\item computed dependencies using non-compositional operational models over
  alternate worlds \cite{Manson:2005:JMM:1047659.1040336,
    DBLP:conf/esop/JagadeesanPR10,
    DBLP:conf/popl/KangHLVD17,DBLP:journals/pacmpl/ChakrabortyV19,DBLP:conf/pldi/LeeCPCHLV20,promising-ldrf}---these
  models validate many compiler optimizations, but also
  %allow some dubious executions 
  fail to validate temporal safety properties
  (see \textsection\ref{sec:promising}).
\end{itemize}

@@

To manage complexity, we have layered the definitions.  After an overview, we
define sequential dependencies in \textsection\ref{sec:model}.  The next two
sections add concurrency.  In \textsection\ref{sec:mca}, we define \PwTmca{},
which provides a Java-like model for \mca{} hardware, similar to that of
\citet{DBLP:journals/pacmpl/JagadeesanJR20}.  In \textsection\ref{sec:c11},
we define \PwTc{}, which models \cXI, adapting the approach of
\citet{DBLP:conf/esop/PaviottiCPWOB20}.  In \textsection\ref{sec:additional},
we extend the semantics to include additional features, such as address
calculation.  In \textsection\ref{sec:results}, we summarize our results.

@@

>>>>

In certain cases it is possible to ban concurrent access using separation
\cite{OHearn:2007:RCL:1235896.1236121}, or to accept inefficient
implementation in order to obtain sequential consistency
\cite{DBLP:conf/snapl/MarinoMMNS15}.  When these approaches are not
available, however, we are left with an enormous gap in our understanding of
one of the most basic elements of computing: the humble semicolon.  Until
recently, existing approaches either
\begin{itemize}
\item did not bother tracking dependencies, allowing ``thin air'' executions
  --- as in C and C++~\cite{DBLP:conf/esop/BattyMNPS15},
\item tracked dependencies conservatively, using syntax, requiring inefficient
  implementation of relaxed access
  \cite{DBLP:conf/pldi/LahavVKHD17,Boehm:2014:OGA:2618128.2618134,DBLP:conf/oopsla/VafeiadisN13,DBLP:journals/corr/abs-1804-04214}--- a non-starter for safe languages like Java,
  and an unacceptable cost for low-level languages like C,
\item computed dependencies using non-compositional operational models over
  alternate worlds \cite{Manson:2005:JMM:1047659.1040336,
    DBLP:conf/esop/JagadeesanPR10,
    DBLP:conf/popl/KangHLVD17,DBLP:journals/pacmpl/ChakrabortyV19,DBLP:conf/pldi/LeeCPCHLV20,promising-ldrf}---these
  models validate many compiler optimizations, but also
  %allow some dubious executions 
  fail to validate temporal safety properties
  (see \textsection\ref{sec:promising}).
\end{itemize}

@@

To manage complexity, we have layered the definitions.  After an overview, we
define sequential dependencies in \textsection\ref{sec:model}.  The next two
sections add concurrency.  In \textsection\ref{sec:mca}, we define \PwTmca{},
which provides a Java-like model for \mca{} hardware, similar to that of
\citet{DBLP:journals/pacmpl/JagadeesanJR20}.  In \textsection\ref{sec:c11},
we define \PwTc{}, which models \cXI, adapting the approach of
\citet{DBLP:conf/esop/PaviottiCPWOB20}.  In \textsection\ref{sec:tool} we 
provide a tool for automatic evaluation of litmus tests.  In \textsection\ref{sec:additional},
we extend the semantics to include additional features, such as address
calculation.  In \textsection\ref{sec:results}, we summarize our results.

@@

Other work in relaxed memory has shown that tooling is especially useful to researchers, architects, and language specifiers, enabling them to build intuitions experimentally~\cite{DBLP:conf/esop/PaviottiCPWOB20,PrideMM,alglave,Batty:2011:MCC:1926385.1926394}.
Unfortunately, it is not obvious that tools can be built for all thin-air free models, the calculation of \citet{Pichon-Pharabod:2016:CSR:2837614.2837616} does not have a termination proof for an arbitrary input, and the enormous state space for the operational models of \citet{DBLP:conf/popl/KangHLVD17} and \citet{DBLP:journals/pacmpl/ChakrabortyV19} is a daunting prospect for a tool builder -- and as yet no tool exists for automatically evaluating these models.
We describe a tool, \PwTer, for automatically evaluating \PwT{} in \textsection\ref{sec:tool}.
